Boutiques: a flexible framework for automated application integration in computing platforms

### 2016-11-11 paper.tex
Updates in related work

@@

The Common Workflow Language is the closest piece of related work as
it supports containers. In particular, the Command Line Tool
Description of CWL overlaps with the \boutiques manifest. According to
Github, CWL started 6 months before Boutiques (September 2014 vs. May
2015). The differences highlighted below were identified from version
1.0 of the CWL Command Line Tool
Description\footnote{\url{http://www.commonwl.org/v1.0/CommandLineTool.html}}.

@@

The following differences are conceptual in the sense that they may
not be easily addressed by CWL or Boutiques developers.

@@

CWL tools have to write in a specific set of directories called
``designated output directory'', ``designated temporary directory''
and ``system temporary directory''. Tools are informed of the location
of such directories through environment variables. Having to write in
specific directories is problematic because tools have to be modified
to enable that. In \boutiques, the path of output files is defined
using a dedicated property.

CWL types are richer, not only
semantically but also syntactically. For instance, files have
properties for basename, dirname, location, path, checksum, etc.

\boutiques supports various types of containers (Docker, Singularity,
rootfs) while CWL supports only Docker. At the same time, CWL’s
requirements are richer: for instance, they may include RAM and disk
while Boutiques is limited to walltime estimate. CWL has hints, i.e.,
recommendations that only lead to warnings when not respected.

In \boutiques, dependencies can be defined among inputs, e.g., to
specify that an input may be used only when a particular flag is
activated. This is a very useful feature to improve validation, in
particular when tools have a lot of options.

In \boutiques, named
groups of inputs can be defined, which improves the presentation of
long parameter lists for the user and enables the definition of more
constraints within groups (e.g. mutually exclusive inputs).

@@

A large number of frameworks have been created to facilitate the
integration of command-line applications in platforms. In
neuroinformatics, several platforms platforms define a formal
interface to embed command-line applications. Among them, the Common
Toolkit\footnote{\url{http://www.commontk.org}} has an
interoperability focus like \boutiques as its command-line modules are
supported by several platforms such as 3D Slicer~\cite{pieper20043d},
NiftyView~\cite{Craddock2016}, GIMIAS~\cite{larrabide2009gimias},
MedInria~\cite{larrabide2009gimias}, MeVisLab~\cite{heckel2009object}
and MITK workbench~\cite{nolden2013medical}. The framework, however,
remains tightly bound to the Common Toolkit's C++ implementation which
limits its adoption, e.g., in web platforms \todo{Fr\'ed\'eric, maybe
  you have more to say on this}.

In the distributed computing community, systems were also proposed to
facilitate the embedding of applications in platforms. The Grid
Execution Management for Legacy Code Architecture
(GEMLCA~\cite{delaitre2005gemlca}) was used to wrap applications in
grid computing systems. Interestingly, it has been used to embed
workflow engines in the SHIWA
platform~\cite{terstyanszky2014enabling}, in a similar but different
way than proposed by \boutiques (Section~\ref{sec:parallelization}).

The recent advent of virtual containers requires a new generation of
application description frameworks that are independent from any
programming language and that expose a rich set of properties to
describe command lines as intended by \boutiques.

>>>>

The Common Workflow Language\footnote{\url{http://www.commonwl.org}}
is the piece of related work most closely related to
\boutiques. Several other application description frameworks have been
developed but they do not support containers.

@@

The Command Line Tool Description of the Common Workflow Language
(CWL) overlaps with the \boutiques manifest. According to the history
of the CWL and \boutiques Github repositories, CWL started 6 months
before Boutiques (September 2014 vs. May 2015). The differences
highlighted below were identified from version 1.0 of the CWL Command
Line Tool
Description\footnote{\url{http://www.commonwl.org/v1.0/CommandLineTool.html}}.

@@

The following differences are conceptual in the sense that they may
not be easily addressed by CWL or Boutiques developers without deeply
refactoring the frameworks.

@@

\begin{itemize}
\item CWL tools have to write in a specific set of directories called
``designated output directory'', ``designated temporary directory''
and ``system temporary directory''. Tools are informed of the location
of such directories through environment variables. Having to write in
specific directories is problematic because tools have to be modified
to enable that. In \boutiques, the path of output files is defined
using a dedicated property.
\item CWL types are richer, not only
semantically but also syntactically. For instance, files have
properties for basename, dirname, location, path, checksum, etc.
\item \boutiques supports various types of containers (Docker, Singularity,
rootfs) while CWL supports only Docker. At the same time, CWL’s
requirements are richer: for instance, they may include RAM and disk
while Boutiques is limited to walltime estimate. CWL has hints, i.e.,
recommendations that only lead to warnings when not respected.
\item In \boutiques, dependencies can be defined among inputs, e.g., to
specify that an input may be used only when a particular flag is
activated. This is a very useful feature to improve validation, in
particular when tools have a lot of options.
\item In \boutiques, named
groups of inputs can be defined, which improves the presentation of
long parameter lists for the user and enables the definition of more
constraints within groups (e.g. mutually exclusive inputs).
\end{itemize}

@@

Several frameworks have been created to facilitate the
integration of command-line applications in platforms. In
neuroinformatics, many platforms define a formal interface to embed
command-line applications. Among them, the Common
Toolkit\footnote{\url{http://www.commontk.org}} interoperates with
several platforms such as 3D Slicer~\cite{pieper20043d},
NiftyView~\cite{Craddock2016}, GIMIAS~\cite{larrabide2009gimias},
MedInria~\cite{larrabide2009gimias}, MeVisLab~\cite{heckel2009object}
and MITK workbench~\cite{nolden2013medical}. The framework, however,
remains tightly bound to the Common Toolkit's C++ implementation which
limits its adoption, e.g., in web platforms \todo{Fr\'ed\'eric, maybe
  you have more to say on this}.

In the distributed computing community, systems were also proposed to
facilitate the embedding of applications in platforms. The Grid
Execution Management for Legacy Code Architecture
(GEMLCA~\cite{delaitre2005gemlca}) was used to wrap applications in
grid computing systems. Interestingly, it has been used to embed
workflow engines in the SHIWA
platform~\cite{terstyanszky2014enabling}, in a similar but different
way than proposed by \boutiques (see Section~\ref{sec:parallelization}).

The recent advent of virtual containers requires a new generation of
application description frameworks that are independent from any
programming language and that expose a rich set of properties to
describe command lines, as intended by \boutiques.
